\chapter{Ondas}
Problemas do Haliday Capítulo 16 e 17

A simple function is given by $y(x) = x( \pi - x)$ in the region
$0 \leqslant x \leqslant \pi$ . It is desired that this function be approximated by
a series of sine functions in the form \[y(x) = a_1  \sin x +
a_3 \sin 3x + a_5 \sin 5x \cdots\] Esse é um exemplo de uma aproximação por série de Fourier. (a) Use a graphing program and
estimate the values for $a_1 , a_3$ , and $a_5$ that give the best visual
fit. (b) Use a symbolic math program (such as Scipy or Scilab) to evaluate the integrals

\[I_n = \int_0^\pi \sin^2 nx \dd x\]
e 
\[I_{nm} = \int_0^\pi \sin nx \sin mx \dd x \]
(c) Find the exact values of the coefficients $a_n$ for
$n \in \{1, 2, 3, 4, 5\}$ by evaluating
\[ a_n = \frac{1}{I_n} \int_0^\pi x(\pi - x) \sin nx \dd x   \]
Why does this work? Compare your answers to the visual in-
spection process.

1. Write a computer program for a Doppler sonar. The program
should request the speed of sound, the frequency of the output
pulse (or “ping”), the frequency of the reflected pulse, and the
time delay between the output ping and the return ping. The
program should then inform the user of the probable distance
to the target and the target’s possible speed(s) toward or away
from the source. Try the program with the following data: the
speed of sound is 340 m/s; the output pulse frequency is
20 kHz; the frequency of the reflected pulse is 20.612 kHz;
and the time delay between the output and reflected pings is
0.230 s.
2. Generalize the previous program so that the data from two
consecutive pings can be used to determine both the distance
to the target and the velocity of the target. The program will
also need to request the rate at which outgoing pings are sent.
Assume that the outgoing pings are omnidirectional, but the
direction of incoming pings can be resolved. Try the program
with the following data: the speed of sound is 340 m/s; the
output pulse frequency is 20 kHz and the pulses are sent once
per second; a 20.921 kHz reflected pulse coming from 40° E
of N is received 0.288 s after the first pulse is sent; and a sec-
ond 20.921 kHz reflected pulse coming from $36.5^o$ E of N is
received 0.311 s after the second pulse is sent.